---
title: Cross Validation"

author: "Mari Sanders"

date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(mgcv)
library(SemiPar)
set.seed(1)
```

# LIDAR Data

```{r}
data("lidar")

lidar_df <- 
  lidar %>% 
  as_tibble() %>% 
  mutate(id = row_number())
```

```{r}
lidar_df %>% 
  ggplot(aes(x = range, y = logratio)) + 
  geom_point()
```

# Cross Validation 

- Compare 3 models, one linear, one smooth, one wiggly 

Construct training and testing df 

```{r}
train_df <- sample_frac(lidar_df, size = 0.8)
test_df <- anti_join(lidar_df, train_df, by = "id")
```

```{r}
ggplot(train_df, aes(x = range, y = logratio)) + 
  geom_point() + geom_point(data = test_df, color = "red")
```

Fit three models 

```{r}
linear_mod <- lm(logratio ~ range, data = train_df)
smooth_mod <- gam(logratio ~ s(range), data = train_df)
wiggly_mod <- gam(logratio ~ s(range, k = 30), sp = 10e-6, data = train_df)
```

Look at fits 

```{r}
train_df %>% 
  add_predictions(linear_mod) %>% 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() + 
  geom_point(data = test_df, color = "red") +
  geom_line(aes(y = pred), color = "red")
```

```{r}
train_df %>% 
  add_predictions(wiggly_mod) %>% 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +  
  geom_point(data = test_df, color = "blue") +
  geom_line(aes(y = pred), color = "red")
```

```{r}
train_df %>% 
  add_predictions(smooth_mod) %>% 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +  geom_point(data = test_df, color = "blue") +
  geom_line(aes(y = pred), color = "red")
```

- Does better than linear model because it is capturing the trends, but not trying to be so flexible like `wiggly_mod`

Compare these numerically using rmse

```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

# Repeat the train/test split

```{r}
cv_df <- 
  crossv_mc(lidar_df, 100) %>% 
  mutate(
    train = map(train, as_tibble), 
    test = map(test, as_tibble)
  )
```


Fit models, extract RMSEs

```{r}
cv_res_df <- 
  cv_df %>% 
  mutate(
    linear_mod = map(train, \(x) lm(logratio ~range, data = x)),
    smooth_mod = map(train, \(x) gam(logratio ~s(range), data = x)),
    wiggly_mod = map(train, \(x) gam(logratio ~s(range, k = 30), sp = 10e-6, data = x))
  ) %>% 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, rmse),
    rmse_smooth = map2_dbl(smooth_mod, test, rmse),
    rmse_wiggly = map2_dbl(wiggly_mod, test, rmse)
  ) 

```

Look at RMSE distribution 

```{r}
cv_res_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(), 
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```

# Nepalese Children Dataframe

```{r}
child_df <- read_csv("data/nepalese_children.csv") %>% 
  mutate(
    weight_ch7 = (weight > 7) * (weight - 7)
  ) 
```

Look at df 

```{r}
child_df %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = 0.5)
```

Fit Models

```{r}
linear_mod = lm(armc ~weight, data = child_df)
pwl_mod = lm(armc ~weight + weight_ch7, data = child_df)
smooth_mod <- gam(armc ~s(weight), data = child_df)

```

Look at models 

```{r}
child_df %>% 
  add_predictions(linear_mod) %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = 0.1) + 
  geom_line(aes(y = pred), color = "red")

child_df %>% 
  add_predictions(pwl_mod) %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = 0.1) + 
  geom_line(aes(y = pred), color = "red")

child_df %>% 
  add_predictions(smooth_mod) %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = 0.1) + 
  geom_line(aes(y = pred), color = "red")
```

CV to select models       
```{r}
cv_df <- 
  crossv_mc(child_df, 100) %>% 
  mutate(
    train = map(train, as_tibble), 
    test = map(test, as_tibble)
  )
```

Apply models and extract RMSE

```{r}
cv_res_df <- 
  cv_df %>% 
  mutate(
    linear_mod = map(train, \(x) lm(armc ~ weight, data = x)), 
    pwl_mod = map(train, \(x) lm(armc ~ weight + weight_ch7, data = x)), 
    smooth_mod = map(train, \(x) gam(armc ~ s(weight), data = x)) 
  ) %>% 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, rmse),
    rmse_pwl = map2_dbl(pwl_mod, test, rmse), 
    rmse_smooth = map2_dbl(smooth_mod, test, rmse)
  )
```

```{r}
cv_res_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(), 
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```

