---
title: "Linear Models"

author: "Mari Sanders"

date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

```{r}
data("nyc_airbnb")

nyc_airbnb <- 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>%  
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighborhood, room_type) %>% 
  mutate(
    borough = fct_infreq(borough), 
    room_type = fct_infreq(room_type)
  )


```

# Fit Models 

Simple Model 

```{r}
fit <- lm(price ~ stars, nyc_airbnb)

summary(fit)
names(summary(fit))

coef(fit)

fit %>% 
broom::tidy() %>% 
  select(term, estimate, p.value)

fit %>% 
  broom::glance()
```

More Complex Model 

```{r}
fit <- 
  lm(price ~ stars + borough, data = nyc_airbnb)

fit %>%  
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(
    term = str_replace(term, "borough", "Borough:")
  ) %>% 
  knitr::kable(digits = 3)
```


# Diagnostics 

EDA
```{r}
nyc_airbnb %>%  
  ggplot(aes(x = stars, y = price)) + 
  geom_point() + 
  stat_smooth(method = "lm")
```

Diagnostics
```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = resid)) + geom_histogram()
```

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = borough, y = resid)) + 
  geom_violin() +
  ylim(-200, 500)
```


Residuals against stars

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = stars, y = resid)) + 
  geom_point()
```

Residuals against fitted values

```{r}
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  modelr::add_predictions(fit) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point()
```

# Hypothesis Testing 

For single coefficients, look at the table 

```{r}
fit %>%  
  broom::tidy() 

fit_null <- lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt <- lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```


For multiple coefficients, look at both 

```{r}
fit_null %>% 
  broom::tidy()

fit_alt %>% 
  broom::tidy()

anova(fit_null, fit_alt) %>% 
  broom::tidy()

```

- Only works for nested models 

# Do Effects Differ Across Boroughs?

First, use a lot of interactions

```{r}
nyc_airbnb %>% 
  lm(price ~ stars * borough + room_type * borough, data = .) %>% 
  broom::tidy()
```

Fit Separate Models 

```{r}
nyc_airbnb %>%  
  filter(borough == "Manhattan") %>% 
  lm(price ~stars + room_type, data = .) %>% 
  broom::tidy()

nyc_airbnb %>%  
  filter(borough == "Brooklyn") %>% 
  lm(price ~stars + room_type, data = .) %>% 
  broom::tidy()
```

Use List Columns 

```{r}
nyc_airbnb %>%  
  nest(data = -borough) %>% 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)), 
    results = map(model, broom::tidy)
    ) %>% 
  select(borough, results) %>% 
  unnest(results) %>% 
  select(borough, term, estimate) %>% 
  pivot_wider(
    names_from  = term, 
    values_from = estimate
  )
```


Room Type Across Manhattan Neighborhoods 

```{r}
nyc_airbnb %>% 
  filter(borough == "Manhattan", neighborhood == "Chinatown") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy()

nyc_airbnb %>% 
  filter(borough == "Manhattan", neighborhood == "Chelsea") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy()
```


For all Neighborhoods 

```{r}
manhattan_neighborhoo_fit_df <- 
nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  nest(data = -(borough:neighborhood)) %>% 
  mutate(model = map(data, \(x) lm(price ~ stars + room_type, data = x)), 
         results = map(model, broom::tidy)) %>% 
  select(neighborhood, results) %>% 
  unnest(results) 
```

```{r}
manhattan_neighborhoo_fit_df %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = term, y = estimate)) + geom_boxplot()
```

```{r}
manhattan_neighborhoo_fit_df %>% 
  filter(str_detect(term, "room_type")) %>% 
  filter(estimate > 0)

nyc_airbnb %>% 
  filter(
    neighborhood == "NoHo", 
    room_type == "Shared room"
  )
```

